---
title: "Missouri air permits"
output: html_notebook
---

The beginning of an analysis of air permit data in Missouri and Kansas City.

Loading in completed projects data set:

```{r}
completed_mo <- read_csv("data/completed-air-permits-july1.csv")
```

At 32,157 columns. 1 column = 1 permit.

### Integrity checks:

How many unique IDs?

```{r}
completed_mo %>% group_by(program_reference_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

32,154 rows/IDs

3 have duplicates of 2 - AP199500016 - AP200110083 - AP200504051

How many installation names?

```{r}
completed_mo %>% group_by(facility_site_name) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

This looks like company name - total of 6,977.

MFA INC. has the most at 256 Capital Quarries, Inc. has 160 Mississippi Lime Company has 137.

How many cities? How many permits per city?

```{r}
completed_mo %>% group_by(city) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

823 cities

St. Louis - 2339 rows Kansas City - 1207 rows St. Joseph - 795 rows Springfield - 720 rows

Also, North Kansas City has 341 rows at #10

How many counties? How many counties per city?

```{r}
completed_mo %>% group_by(county) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

Jackson County - 1942 St. Louis City - 1842 St. Louis - 1180 St. Charles - 1038 Jasper - 988 Portable Plant - 974

How many permit statuses?

```{r}
completed_mo %>% group_by(permit_status) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

55 statuses

AP: No permit required is the most at 4883 AP: Permit Issued is second at 4063 This is a dirty column\*\*

How many permit types?

```{r}
completed_mo %>% group_by(permit_type) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

53 permit types

How many permit numbers?

```{r}
completed_mo %>% group_by(permit_number) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

13,822 permit numbers (8,424 NAs, 127 labeled as N/A - column might not be the most reliable).

How many site ids?

```{r}
completed_mo %>% group_by(site_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

7,983 site IDs

Filtering by Kansas City

```{r}
kc_completed <- completed_mo %>% filter(city == "Kansas City")
```

1207 rows / permits

### In Kansas City:

What company has the most applications?

```{r}
kc_completed %>% group_by(facility_site_name) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

Ford Motor Company - KC Plant at 51 Miles Inc at 39 Kansas City Power and Light (Hawthorn) has 37 (third highest)

What type of permit is the most?

```{r}
kc_completed %>% group_by(permit_type) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

AP: Local CP - 311 AOP: Basic operating permit renewal - 191

Mapping Where are there permits issues in Kansas City?

How many KC rows are missing an address?

```{r}
kc_completed %>% group_by(facility_site_address) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

11 blanks. This also shows 51 permits at the same location, and then a lot at the Hawthorn location (Hawthorn came up in research).

Making addresses lowercase, then geocoding and loading back in:

```{r}
kc_completed <- kc_completed %>% mutate(facility_site_address = tolower(facility_site_address))

write_csv(kc_completed, "data/kc-completed-july1.csv")

kc_geo <- read_csv("data/kc-completed-july1-geocoded.csv") %>% clean_names()
```

Mapping the geocoded data as points on a map to look for trends:

```{r}
leaflet(kc_geo) %>% addTiles() %>% addCircles(lng = ~longitude, lat = ~latitude)
```

What census tract has the most?

```{r}
kc_geo %>% group_by(full_fips_tract) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

29095015500 (155) has 192 permits

175 has 68

221 has 65
